Rotational Spectral Unmixing of Exoplanets:\\Degeneracies between Surface Colors and Geography

### 2017-04-28 paper/main.tex
final proofreading

%%% http://www.oceanwave.jp/index.php?float%B4%C4%B6%AD(figure%2Ftable)%A4%CE%BD%D0%CE%CF%B0%CC%C3%D6%A4%F2%A5%B3%A5%F3%A5%C8%A5%ED%A1%BC%A5%EB

@@

Unmixing the disk-integrated spectra of exoplanets provides a clue to heterogeneous surfaces that we cannot directly resolve in the foreseeable future. 
It is particularly important for terrestrial planets with diverse surface compositions like Earth. 
Although previous work on unmixing the spectra of Earth from disk-integrated multi-band light curves appeared successful, we point out a mathematical degeneracy between the surface colors and their spatial  distributions. 
Nevertheless, useful constraints on the spectral shape of individual surface types may be obtained from the fact that albedo should be between 0 and 1, depending on the trajectory of the planet in color space.  
We demonstrate the degeneracy and the possible constraints using both mock data based on a toy model of Earth, as well as observations of Earth. 
Despite the severe degeneracy, we are still able to recover an approximate albedo spectra for ocean. 
In general, we find that surfaces are easier to identify when they cover a large fraction of the planet and when their spectra approach zero or unity in certain bands. 
%We find that observing at multiple wavelengths where surfaces albedo is close to 0 or 1 will provide better constraints. 
\end{abstract}

@@

Direct imaging of exoplanets is expected to play a vital role in characterizing Earth analogs in habitable zones and beyond. 
Substantial work has gone into prediciting detectable features in disk-integrated spectra of Earth and other planets, as they are observed from an astronomical distance. 
Disk-integrated spectra or even multi-band photometry (colors) in principle include the effects of the surface reflectance as well as the atmosphere. (We use the term spectra and colors interchangeably.) 
%Atmospheric molecules are identifiable through spectral absorption features \citep[e.g.,][]{DesMarais2002} but surface reflectance spectra also affect the spectra, which could be measured through low-resolution spectra, or multi-band photometry. 
However, interpreting disk-integrated colors is not trivial. 
This is particularly true for Earth-like planets that harbor diverse atmospheric and surface characteristics including liquid water, partial cloud cover, continents, and possibly vegetation.

A key here is to leverage the time variation of the spectrum/color  \citep{Ford2001}: the regions that contributes to the scattered light change due to the planetary rotation and  orbital motion, so the time variability can be used to map the heterogeneity of the surface environment \citep[see][for recent review]{Cowan2017}.  
\citet{Cowan2009, Cowan2011} performed Principal Component Analysis (PCA) on the observed multi-band photometry of Earth. They found that the number of surface types can be inferred from the number of dominant principle components: (\# of surface types) $\ge $ (\# of principal components) $+ 1$. %) and that the variation pattern is indicative of the surface properties. 
\citet{Fujii2010, Fujii2011} decomposed multi-band photometric data of the Earth assuming the template reflectance spectra of the known major surface types, and they found that the relative abundance and longitudinal variation of these surfaces are approximately recovered. 
Moreover, by coupling the time variation due to rotation with phase variation due to orbital revolution, a 2-dimensional map of the surface may be retrieved \citep{Kawahara2010, Kawahara2011, Fujii2012}.

@@

Motivated by this unsatisfactory result, we revisited their analysis. 
We start by creating mock light curves to see if we can fully retrieve the known answer. 
Taking a close look at the mathematical relation between the colors and geographical maps of the surfaces, we discuss the degeneracy among them. 
We then demonstrate using mock data how our estimates are constrained by boundary conditions, including the expectation that albedo should be between 0 and 1. 
We also apply the same procedure to the data observed by EPOXI.

@@

The assumed albedo spectra of these surface types in 4 photometric bands are shown in the upper right panel of Figure \ref{fig:mockdata}. 
These 4 photometric bands correspond to 0.4--0.5 $\mu $m, 0.5--0.6 $\mu $m, 0.6--0.7 $\mu $m, and 0.7--0.8 $\mu $m, respectively, but we will simply refer to them by the band indices unless otherwise noted. 
%\memoJLY{I think it would be helpful to make the "band \#"-axis wavelength in microns, and then perhaps refer to the band \# in the text or in an upper x-axis. }
The albedo spectrum for ocean is based on \citet{Mclinden1997}, 
and the data for sand and vegetation are taken from ASTER spectral library\footnote{\url{https://speclib.jpl.nasa.gov/}. 
Specifically, we adopt  ``Brown to dark brown sand'' for ``sand'', and ``Grass'' for the ``vegetation''}. 
The scattering phase function by the surface is assumed to obey the Lambert law, i.e., the radiance is (incident flux)$\times $(albedo)/$\pi$, independent of the direction of the scattering. 
Note that in reality they are not Lambertian scatterers; among others,  scattering by the ocean is particularly anisotropic at crescent phase (see Section \ref{ss:deviate_Lambert}).

% and the albedo spectra of ocean overlaid with atmosphere  
%\memoJLY{A phase of 135 deg is right in the range that Ty (Robinson et al. 2010; 2011) shows will deviate most from a Lambertian scatterer due the ocean glint...}

%{\color{red} 
The diurnal light curves are synthesized given the relative positions of the star, planet, and observer. 
% We consider light curves of 1 rotation at a fixed orbital phases. 
For the sake of simplicity, we consider a planet with zero obliquity in an edge-on orbit, and change the phase angle (the planet-centric angle between the star and the observer) denoted by $\alpha $. 
We also assume that spin period is significantly shorter than the orbital period and the orbital phase does not change in a single spin rotation. 
However, the following discussion does not depend on these assumptions. 
We consider noise-free data. Clearly, noise in real observations of Earth twins are expected to be substantial, and the effect of such observations will be discussed elsewhere. 
%\color{black}

The bottom panels of Figure \ref{fig:mockdata} display examples of diurnal light curves in 4 photometric bands, represented in terms of {\it apparent albedo}. 
Apparent albedo is the albedo of the planet weighted by illumination and visibility; it is obtained by normalizing the planetary intensity by that of a loss-less Lambert sphere with the same radius and at the same phase \citep{Qiu2003, Seager2010}. 
In this paper we use apparent albedo unless otherwise noted. 
The apparent albedo is straightforwardly obtained from the observed planetary intensity, if and only if the planetary radius, orbital phase of the planet, and the distance between the star and the planet are all known.

@@

In this section, we discuss the general framework for analying the diurnal light curves, and present some demonstrations using mock data created in the previous section. 
Sections \ref{ss:model} and \ref{ss:PCplane} are essentially the recapitulation of previous papers, in particular \citet{Cowan2013} \citep[but see also][]{Cowan2009,Cowan2011,Fujii2010,Fujii2011}.  
We choose to include these discussions, however, as a baseline to establish the later arguments.

@@

Assuming that the planetary surface is everywhere Lambertian scatterering, and that it is composed of a certain number $K$ of spectrally distinct surface types, the disk-integrated scattered light is a weighted summation of the reflectance spectra of different surface types. 
Using the local surface albedo $s_{\vec \Omega }$, the zenith angle of the insolation, $\theta _0$, and the zenith angle of the observer $\theta _1$ (both defined at each surface point),
the apparent albedo of the planet in band $j$ at epoch $i$, $d_{ij}$ (``$d$'' for data) is written as follows \citep[see][]{Fujii2010}: 
%%%
\begin{eqnarray}
d_{i} (\lambda_j) &=& \displaystyle \frac{ \int_{{\rm IV}_i} s_{\vec \Omega }(\lambda_j) \cdot \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega } \\
&=& \sum _{k} s_k (\lambda_j) \; \displaystyle \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega } \notag \\
&=& \sum _{i,k} \fast_{ik} \, s_{kj} \label{eq:tilde_d_f_ast_s}
\end{eqnarray}
%%%
where the spectrum of $k$-th surface type is represented by:
%%%
\begin{equation}
s _{kj} \equiv  s_k (\lambda _j)
\end{equation}
%%%
and the apparent covering fraction of $k$-th surface type at epoch $i$ is:
%%%
\begin{equation}
\tilde f_{ik} \equiv  \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }
\end{equation}
%%%
% and $i$, $j$, and $k$ are indices for the observation epochs, bands, and the surface types, respectively, 
where ${\rm IV }_i$ denotes the illuminated and visible area over the planetary surface at epoch $i$, and $f_k (\vec \Omega )$ is the area fraction of $k$-th surface type in $d\vec \Omega$. 
%In the last expression, $\fast_{ik}$ represents the apparent covering fraction of the $k$-th surface type at $i$-th observational epoch, and 
%$s_{kj}$ is the reflectance spectra of the $k$-th surface type in the $j$-th band. 
The maximum value of $i$, $j$ and $k$ will be denoted by $I$, $J$, and $K$, below, as summarized in Table \ref{tab:index}.

@@

%%%
%\begin{eqnarray}
%\begin{cases}
%\;\; 0 \leq s_{kj} \leq 1 \;\;\; & \mbox{for any $k$, $j$} \\
%\;\; 0 \leq \fast_{lk} \leq 1 \;\;\; & \mbox{for any $l$, $k$} \label{eq:cond_f_ast}\\
%\;\; \sum_k \fast_{lk} = 1 & \mbox{for any $l$} 
%\end{cases}
%\end{eqnarray}
%%%

%\subsection{Estimating the Composition\\of Longitudinal Slices}

The time variability of the apparent covering fraction $\fast $ due to the planet's rotation is related to the surface inhomogeneity in the longitudinal direction. 
Approximately, $\fast _{ik}$ may be written as the weighted summation of the area fraction of the $k$-th surface type in each longitudinal slice, i.e.,
%%%
\begin{eqnarray}
\fast _{ik} &=& \frac{ \int_{{\rm IV}_i} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }  \\
&=& \sum_l \frac{ \int f_k (\vec \Omega ) W_i (\vec \Omega  ) \cdot d\vec \Omega }{ \int  W_i (\vec \Omega ) \cdot d\vec \Omega } \\
%&\approx & \sum_l f_{lk} \frac{ \int W_i (\vec \Omega ) \cdot d\vec \Omega }{ \int  W_i (\vec \Omega ) \cdot d\vec \Omega } \label{eq:discretize}\\
&\approx & \sum_l  W_{il} f_{lk} \label{eq:Wf}, \\
W_{il} &\equiv & \frac{ \int  W_i (\vec \Omega  ) \cdot d\vec \Omega_l }{ \int W_i (\vec \Omega )  \cdot d\vec \Omega }
\end{eqnarray}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type in the $l$-th longitudinal slice. 
Strictly speaking, the approximation is valid only when $f_k(\vec \Omega)$ does not change or changes little across the $l$-th slice for all $k$ and $l$. 
In the last equation, $W_{il}$ is the weight of the $l$-th longitudinal slice at epoch $i$, which depends only on the observational geometry. 
\citet{Cowan2013} refer to this as the convolution kernel, $K$. 
% As a result,

@@

%A question is whether the above formulation leads us to a unique solution of either $\{ {\bf \fast},\,{\bf s} \}$ or $\{ {\bf f},\,{\bf s}\}$, given the data matrix, ${\bf d}$.

Equation (\ref{eq:tilde_d_f_ast_s}) coupled with the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}) indicates a geometrical relationship among ${\bf d}$, ${\bf \fast }$, and ${\bf s}$ in 
$J$-dimensional space, where $\{{\bf d}_i\}$ are located on the plane defined by $K$ surfaces, $\{{\bf s}_k\} $, and in the region enclosed by $\{{\bf s}_k\} $. 
%\memoYF{How is it called in one word in mathematics?}
Figure \ref{fig:schematic} graphically shows this relations in the case of $J=3$ and $K=3$. 
Note that the dimension of this plane is $K-1$, so in general it is a hyper-plane.  
%We show this graphically in Figure \ref{fig:trajectory}, using the mock datasets shown in Section \ref{s:mockdata}.

@@

Using $V_{nj}$ and the time average of the colors in the case of $\alpha = 90^{\circ }$, $\bar d_j$, the light curves are projected onto the PC plane by:
%%%
\begin{equation}
d_{ij} = \sum_n U_{in} V_{nj} + \bar d_j
\end{equation}
%%%
where $U_{in}$ is the trajectory. %\edit{\sout{The $V_{nj}$) and the time-averaged colors ($\bar d_j$) uniquely specify the PC plane. }
% in the PC plane and $\bar d_j$ is the time average of the colors. % as the offset. 
The trajectory of the light curves at different orbital phases are shown by the olive line in Figure \ref{fig:trajectory}. 
%\edit{\sout{The dominant PCs ($V_{nj}$) and the time-averaged colors ($\bar d_j$) uniquely specify the PC plane. }}
%We may plot the trajectories of other light curves at different phase angles on the same plane because the PC plane should be the same\edit{\sout{ plane regardless of orbital phase}}. %(under the noiseless condition) because we use the same 3 surface types and assume Lambertian surfaces. 
% (Note that PCA on individual light curves does not necessarily find the same PCs, because it depends on how the data points are scattered on the PC plane.)
%Thus, the trajectories of the mock light curves at different phase angles are also shown in the lower panel of Figure \ref{fig:trajectory}. 
The color excursions are greater at a larger phase angle (crescent phase), because less surface area is averaged together.  
The circle, square, and triangle in the figure indicate the input albedo spectra of the three surface types. 
As described above, the trajectories are always inside the triangle defined by these points. 
%Because the apparent albedo at any time of the observations is a linear combination of input surface colors (Equation \ref{eq:tilde_d_f_ast_s}) under the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}), the trajectories should always be inside the triangle defined by these points.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}[tbh!]
%    \begin{center}
%\includegraphics[width=\hsize]{PCA_V_jn.pdf}
%    \end{center}
%    \caption{Principal components of the light curves shown in the bottom of Figure \ref{fig:mockdata}. }
%\label{fig:PCs}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%Although this is not sufficient to result in a unique solution, it does impact the precision of rotational spectral unmixing as discussed below.

@@

A formally equivalent degeneracy is found in Equation (\ref{eq:d_f_s}) coupled with the conditions (\ref{eq:f_range})-(\ref{eq:s_range}). 
Essentially, the term $\sum _k f_{lk} s_{kj}$ represents the average albedo spectra of $l$-th slice and $W_{il}$ is the matrix that convolve them into the light curves. 
Suppose idealistically that we can retrieve the average spectra of longitudinal slices from the light curves through $W_{il}$; the black trajectory in the bottom panel of Figure \ref{fig:trajectory} represents the color variation as a function of longitude.  
We again have different sets of $\{ {\bf s}_k \}$ that enclose the averaged albedo spectra of longitudinal slices {\it and} are located in the range of condition (\ref{eq:s_range}). 
%, any of which can make up the given averaged albedo spectra with the associated $f_{lk}$. 
Nevertheless, the excursions of the longitudinal colors are more dramatic than the disk-integrated colors of the light curves, thus the possible solutions are somewhat restricted. 
% For disk-integrated light curves, the color excursions are more muted than the intrinsic longitudinal map, so the degeneracy becomes severer.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbh!]
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t360_lc_noreg_allowedregion_gray.pdf}
%\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_noreg.pdf}
    \end{center}
     \end{minipage}   
    \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_135deg_3types_t360_lc_noreg_allowedregion_gray.pdf}
%\includegraphics[width=\hsize]{mockdata_135deg_3types_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
%\includegraphics[width=\hsize]{IGBP_lon_noreg.pdf}
\includegraphics[width=\hsize]{IGBP_lon_noreg_allowedregion_gray.pdf}
    \end{center}
     \end{minipage}
    \caption{Probability of surface albedo spectra estimated from the light curves of $\alpha = 90^{\circ }$ (left), from the light curves of $\alpha = 135^{\circ }$ (middle) and from the colors of longitudinal slices. }
\label{fig:PCplane}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Given that any three vertices in the principal component plane that enclose the trajectory/-ies can be a solution, the choice of the solution depends on the prior probability distribution. 
% Thus, we need to be careful about the implicit assumptions that are made  
With no assumptions or information on the spectral albedo of the surface types or geography, it is reasonable to assume that any points in the PC plane except for the inhibited regions are equally likely to correspond to a surface type. 
Despite this uninformative prior, the location of the light curve trajectory allows us to constrain the location of the surface spectra in a probabilistic sense. 
%While such an assumption does not seem to work at all to choose the surface types, it does put some constrains. 
%While such an assumption does not uniquely specify the colors of all $K$ surface types, it does offer constraints due to the relative light curve trajectory within the permitted region. 
%This is because of the relative configuration of the trajectory and the permitted region. 
For example, in Figure \ref{fig:trajectory}, in order to enclose the trajectory with three surface types within the permitted region, we must have at least one point near the bottom left corner of the permitted region; this is consistent with the fact that one of the input albedo spectra (ocean) resides there.

@@

Meanwhile, in (b) we swap sand and ocean.  
Although sand is now the dominant component, 
%the true location of the soil is rather not preferred based on 
our analysis prefers other locations toward the top of the allowed region, because there is more space there. 
Therefore, large covering fraction is not a sufficient condition for obtaining strong constraints on the color of a surface. 
Better constraints on a surface spectrum can be obtained if its location on the PC plane is close to more than one boundary of the allowed region. 
In other words, a surface spectrum can be well retrieved if it has a favorable geography (ideally an entire hemisphere solely covered with that surface) {\it and} it has favorable albedo spectrum (albedo close to 0 or 1 at as many wavelengths as possible).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
	\includegraphics[width=0.85\hsize]{PCs_raddata_2_norm.eps}
	\includegraphics[width=0.88\hsize]{raddata_2_norm_noreg_ver2.pdf}
%	\includegraphics[width=\hsize]{raddata_2_norm_Mj_PCs.pdf}
	\includegraphics[width=0.85\hsize]{raddata_2_norm_spectra.pdf}
    \end{center}
    \caption{Top panel: PCs from EPOXI June data. Middle panel: color contour similar to Figure \ref{fig:PCplane}, but based on EPOXI observation of June 2008 \citep{Livengood2011}. The red line shows the light curve trajectory. Bottom panel: spectra corresponding to the locations (A$\sim$J) shown in the upper panel. The ocean, sand, and clouds appear to be close to A, D, and F, respectively. }
\label{fig:EPOXI}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

@@

Unmixing the disk-integrated spectra of exoplanets provides a clue to heterogeneous surfaces that we cannot directly resolve in the foreseeable future. 
It is particularly important for terrestrial planets with diverse surface compositions like Earth. 
Although previous work on unmixing the spectra of Earth from disk-integrated multi-band light curves appeared successful, we point out a mathematical degeneracy between the surface colors and their spatial  distributions. 
Nevertheless, useful constraints on the spectral shape of individual surface types may be obtained from the premise that albedo is between 0 and 1, depending on the trajectory of the planet in color space.  
We demonstrate the degeneracy and the possible constraints using both mock data based on a toy model of Earth, as well as observations of Earth. 
Despite the severe degeneracy, we are still able to recover an approximate albedo spectra for ocean. 
In general, we find that surfaces are easier to identify when they cover a large fraction of the planet and when their spectra approach zero or unity in certain bands. 
\end{abstract}

@@

Direct imaging of exoplanets is expected to play a vital role in characterizing Earth analogs in habitable zones and beyond. 
Substantial work has gone into prediciting detectable features in disk-integrated spectra of Earth and other planets, as they are observed from an astronomical distance. 
Disk-integrated spectra or even multi-band photometry (colors) in principle include the effects of the surface reflectance as well as the atmosphere. (We use the term spectra and colors interchangeably.) 
However, interpreting disk-integrated colors is not trivial. 
This is particularly true for Earth-like planets that harbor diverse atmospheric and surface characteristics including liquid water, partial cloud cover, continents, and possibly vegetation.

A key here is to leverage the time variation of the spectrum/color  \citep{Ford2001}: the regions that contributes to the scattered light change due to the planetary rotation and  orbital motion, so the time variability can be used to map the heterogeneity of the surface environment \citep[see][for recent review]{Cowan2017}.  
\citet{Cowan2009, Cowan2011} performed Principal Component Analysis (PCA) on the observed multi-band photometry of Earth. They found that the number of surface types can be inferred from the number of dominant principle components: (\# of surface types) $\ge $ (\# of principal components) $+ 1$.  
\citet{Fujii2010, Fujii2011} decomposed multi-band photometric data of the Earth assuming the template reflectance spectra of the known major surface types, and they found that the relative abundance and longitudinal variation of these surfaces are approximately recovered. 
Moreover, by coupling the time variation due to rotation with phase variation due to orbital revolution, a 2-dimensional map of the surface may be retrieved \citep{Kawahara2010, Kawahara2011, Fujii2012}.

@@

Motivated by this unsatisfactory result, we revisited their analysis. 
We start by creating mock light curves to see if we can fully retrieve the known answer. 
Taking a close look at the mathematical relation between the colors and geographical maps of the surfaces, we discuss the degeneracy among them. 
We then demonstrate using mock data how our estimates are nevertheless constrained by boundary conditions, including the expectation that albedo is between 0 and 1. 
We also apply the same procedure to the data observed by EPOXI.

@@

The assumed albedo spectra of these surface types in 4 photometric bands are shown in the upper right panel of Figure \ref{fig:mockdata}. 
These 4 photometric bands correspond to 0.4--0.5 $\mu $m, 0.5--0.6 $\mu $m, 0.6--0.7 $\mu $m, and 0.7--0.8 $\mu $m, respectively, but we will simply refer to them by the band indices ($j$) unless otherwise noted. 
The albedo spectrum for ocean is based on \citet{Mclinden1997}, 
and the data for sand and vegetation are taken from ASTER spectral library\footnote{\url{https://speclib.jpl.nasa.gov/}. 
Specifically, we adopt  ``Brown to dark brown sand'' for ``sand'', and ``Grass'' for the ``vegetation''}. 
The scattering phase function by the surface is assumed to obey the Lambert law, i.e., the radiance is (incident flux)$\times $(albedo)/$\pi$, independent of the direction of the scattering. 
Note that in reality they are not Lambertian scatterers; among others,  scattering by the ocean is particularly anisotropic at crescent phase (see Section \ref{ss:deviate_Lambert}).

The diurnal light curves are synthesized given the relative positions of the star, planet, and observer. 
For the sake of simplicity, we consider a planet with zero obliquity in an edge-on orbit, and change the phase angle (the planet-centric angle between the star and the observer) denoted by $\alpha $ to $45^{\circ }$, $90^{\circ }$, and $135^{\circ }$. 
We also assume that spin period is significantly shorter than the orbital period and the orbital phase does not change in a single spin rotation. 
However, the following discussion does not depend on these assumptions. 
We consider noise-free data. Clearly, noise in real observations of Earth twins are expected to be substantial, and the effect of such observations will be discussed elsewhere.

The bottom panels of Figure \ref{fig:mockdata} display examples of diurnal light curves in 4 photometric bands, represented in terms of {\it apparent albedo}. 
Apparent albedo is the albedo of the planet weighted by illumination and visibility; it is obtained by normalizing the planetary intensity by that of a loss-less Lambert sphere with the same radius and at the same phase \citep{Qiu2003, Seager2010}. 
In this paper we use apparent albedo unless otherwise noted. 
The apparent albedo is straightforwardly obtained from the observed planetary intensity, if and only if the orbital phase of the planet, the distance between the star and the planet, and the planetary radius are all known (Section \ref{sss:uncertain_radius}).

@@

In this section, we discuss the general framework for analying the diurnal light curves, and present some demonstrations using mock data created in the previous section. 
Sections \ref{ss:model} and \ref{ss:PCplane} are essentially the recapitulation of previous papers, in particular \citet{Cowan2013} \citep[see also][]{Cowan2009,Cowan2011,Fujii2010,Fujii2011}.  
We choose to include these discussions, however, as a baseline to establish the later arguments.

@@

Assuming that the planetary surface is everywhere Lambertian scatterering, and that it is composed of a certain number $K$ of spectrally distinct surface types, the disk-integrated scattered light is a weighted summation of the reflectance spectra of different surface types. 
Using the local surface albedo $s_{\vec \Omega }$, the zenith angle of the insolation, $\theta _0$, and the zenith angle of the observer $\theta _1$ (both defined at each surface point),
the apparent albedo of the planet in band $j$ at epoch $i$, $d_{ij}$ (``$d$'' for data) is written as follows: 
%%%
\begin{eqnarray}
d_{i} (\lambda_j) &=& \displaystyle \frac{ \int_{{\rm IV}_i} s_{\vec \Omega }(\lambda_j) \cdot \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega } \\
&=& \sum _{k} s_k (\lambda_j) \; \displaystyle \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega } \notag \\
&=& \sum _{i,k} \fast_{ik} \, s_{kj} \label{eq:tilde_d_f_ast_s}
\end{eqnarray}
%%%
where the spectrum of $k$-th surface type is represented by:
%%%
\begin{equation}
s _{kj} \equiv  s_k (\lambda _j)
\end{equation}
%%%
and the apparent covering fraction of $k$-th surface type at epoch $i$ is:
%%%
\begin{equation}
\tilde f_{ik} \equiv  \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }
\end{equation}
%%%
where ${\rm IV }_i$ denotes the illuminated and visible area over the planetary surface at epoch $i$, and $f_k (\vec \Omega )$ is the area fraction of $k$-th surface type in $d\vec \Omega$. 
The maximum value of $i$, $j$ and $k$ will be denoted by $I$, $J$, and $K$, below, as summarized in Table \ref{tab:index}.

@@

The time variability of the apparent covering fraction $\fast $ due to the planet's rotation is related to the surface inhomogeneity in the longitudinal direction. 
Approximately, $\fast _{ik}$ may be written as the weighted summation of the area fraction of the $k$-th surface type in each longitudinal slice, i.e.,
%%%
\begin{eqnarray}
\fast _{ik} &=& \frac{ \int_{{\rm IV}_i} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }  \\
&=& \sum_l \frac{ \int f_k (\vec \Omega ) W_i (\vec \Omega  ) \cdot d\vec \Omega }{ \int  W_i (\vec \Omega ) \cdot d\vec \Omega } \\
&\approx & \sum_l  W_{il} f_{lk} \label{eq:Wf}, \\
W_{il} &\equiv & \frac{ \int  W_i (\vec \Omega  ) \cdot d\vec \Omega_l }{ \int W_i (\vec \Omega )  \cdot d\vec \Omega }
\end{eqnarray}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type in the $l$-th longitudinal slice. 
Strictly speaking, the approximation is valid only when $f_k(\vec \Omega)$ does not change or changes little across the $l$-th slice for all $k$ and $l$. 
In the last equation, $W_{il}$ is the weight of the $l$-th longitudinal slice at epoch $i$, which depends only on the observational geometry. 
\citet{Cowan2013} refer to this as the convolution kernel, $K$.

@@

Equation (\ref{eq:tilde_d_f_ast_s}) coupled with the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}) indicates a geometrical relationship among ${\bf d}$, ${\bf \fast }$, and ${\bf s}$ in 
$J$-dimensional space, where $\{{\bf d}_i\}$ are located on the plane defined by $K$ surfaces, $\{{\bf s}_k\} $, and in the region enclosed by $\{{\bf s}_k\} $. 
Figure \ref{fig:schematic} graphically shows this relations in the case of $J=3$ and $K=3$. 
Note that the dimension of this plane is $K-1$, so in general it is a hyper-plane.

@@

Using $V_{nj}$ and the time average of the colors in the case of $\alpha = 90^{\circ }$, $\bar d_j$, the light curves are projected onto the PC plane by:
%%%
\begin{equation}
d_{ij} = \sum_n U_{in} V_{nj} + \bar d_j
\end{equation}
%%%
where $U_{in}$ is the trajectory on the plane. 
The trajectory of the light curves at different orbital phases are shown by the olive line in Figure \ref{fig:trajectory}. 
The color excursions are greater at a larger phase angle (crescent phase), because less surface area is averaged together.  
The circle, square, and triangle in the figure indicate the input albedo spectra of the three surface types. 
As described above, the trajectories are always inside the triangle defined by these points.

@@

@@

A formally equivalent degeneracy is found in Equation (\ref{eq:d_f_s}) coupled with the conditions (\ref{eq:f_range})-(\ref{eq:s_range}). 
Essentially, the term $\sum _k f_{lk} s_{kj}$ represents the average albedo spectra of $l$-th slice and $W_{il}$ is the matrix that convolve them into the light curves. 
Suppose idealistically that we can retrieve the average spectra of longitudinal slices from the light curves through $W_{il}$; the black trajectory in the bottom panel of Figure \ref{fig:trajectory} represents the color variation as a function of longitude.  
We again have different sets of $\{ {\bf s}_k \}$ that enclose the averaged albedo spectra of longitudinal slices {\it and} are located in the range of condition (\ref{eq:s_range}). 
Nevertheless, the excursions of the longitudinal colors are more dramatic than the disk-integrated colors of the light curves, thus the possible solutions are somewhat restricted.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbh!]
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t360_lc_noreg_allowedregion_gray.pdf}
    \end{center}
     \end{minipage}   
    \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_135deg_3types_t360_lc_noreg_allowedregion_gray.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_lon_noreg_allowedregion_gray.pdf}
    \end{center}
     \end{minipage}
    \caption{Probability of surface albedo spectra estimated from the light curves of $\alpha = 90^{\circ }$ (left), from the light curves of $\alpha = 135^{\circ }$ (middle) and from the colors of longitudinal slices. }
\label{fig:PCplane}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Given that any three vertices in the principal component plane that enclose the trajectory/-ies can be a solution, the choice of the solution depends on the prior probability distribution.  
With no assumptions or information on the spectral albedo of the surface types or geography, it is reasonable to assume that any points in the PC plane except for the inhibited regions are equally likely to correspond to a surface type. 
Despite this uninformative prior, the location of the light curve trajectory allows us to constrain the location of the surface spectra in a probabilistic sense. 
For example, in Figure \ref{fig:trajectory}, in order to enclose the trajectory with three surface types within the permitted region, we must have at least one point near the bottom left corner of the permitted region; this is consistent with the fact that one of the input albedo spectra (ocean) resides there.

@@

Meanwhile, in (b) we swap sand and ocean.  
Although sand is now the dominant component, 
our analysis prefers other locations toward the top of the allowed region, because there is more space there. 
Therefore, large covering fraction is not a sufficient condition for obtaining strong constraints on the color of a surface. 
Better constraints on a surface spectrum can be obtained if its location on the PC plane is close to more than one boundary of the allowed region. 
In other words, a surface spectrum can be well retrieved if it has a favorable geography (ideally an entire hemisphere solely covered with that surface) {\it and} it has favorable albedo spectrum (albedo close to 0 or 1 at as many wavelengths as possible).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
	\includegraphics[width=0.85\hsize]{PCs_raddata_2_norm.eps}
	\includegraphics[width=0.88\hsize]{raddata_2_norm_noreg_ver2.pdf}
	\includegraphics[width=0.85\hsize]{raddata_2_norm_spectra.pdf}
    \end{center}
    \caption{Top panel: PCs from EPOXI June data. Middle panel: color contour similar to Figure \ref{fig:PCplane}, but based on EPOXI observation of June 2008 \citep{Livengood2011}. The red line shows the light curve trajectory. Bottom panel: spectra corresponding to the locations (A$\sim$J) shown in the upper panel. The ocean, sand, and clouds appear to be close to A, D, and F, respectively. }
\label{fig:EPOXI}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
